### 2008-07-03 trunk/Papers/rmrs-08/abstract.tex
3rd July

\begin{abstract}
  One way to construct semantic
  representations in a robust manner is to enhance shallow language
  processors with semantic components (e.g., \cite{frank:2004,
    briscoe:etal:2006}).  Here, we provide a model theory for a
  semantic formalism 
  that is designed to enable this,
  namely Robust Minimal Recursion Semantics ({\sc rmrs},
  \cite{copestake:2003}).  We show that {\sc rmrs} supports a notion
  of entailment 
  that has a number of desirable properties: specifically, it allows
  {\sc rmrs} to form the basis for comparing the semantic output of
  different parses of the same string, even when the parses vary in
  the depth of their analysis, and in particular to test the
  compatibility of the semantic output of a shallow parser with that
  of a deep parser (where such deep parses exist).
\end{abstract}

>>>>

\begin{abstract}
  One way to construct semantic
  representations in a robust manner is to enhance shallow language
  processors with semantic components (e.g.,
  \cite{frank:2004,briscoe:etal:2006}).  Here, we provide a model
  theory for a 
  semantic formalism 
  that is designed to enable this,
  namely Robust Minimal Recursion Semantics ({\sc rmrs},
  \cite{copestake:2003}).  We show that {\sc rmrs} supports a notion
  of entailment 
  that has a number of desirable properties: specifically, it allows
  {\sc rmrs} to form the basis for comparing the semantic output of
  different parses of the same string, even when the parses vary in
  the depth of their analysis, and in particular to test the
  compatibility of the semantic output of a shallow parser with that
  of a deep parser (where such deep parses exist).
\end{abstract}

### 2008-07-03 trunk/Papers/rmrs-08/introduction.tex
3rd July

Representing semantics as a logical form that supports automated
inference and model construction is vital for deeper language
engineering tasks, such as dialogue systems.  Such detailed semantic
representations can be obtained from hand-crafted deep grammars (e.g.,
\cite{butt:etal:1999,copestake:flickinger:2000}), but deep parsing
tends to lack robustness: hand-crafted grammars fail to cover all
words and linguistic constructions; and unedited text and speech
contains ill-formed phrases, which by design deep grammars do not
handle.

@@

However, there are certain semantic
phenomena that such shallow
approaches can't capture reliably, including quantifier scope, optional
arguments,
% COMMENT: this issue about optional arguments is really too cryptic.
% Basically, there is no way for assessing by *any* learned language
% processor whether a single word form that appears in
% a number of different syntactic valencies is an example of lexical
% sense ambiguity, or they are all examples of the same sense being
% used in syntactically different ways (i.e., we have optional
% arguments).  Only hand-crafted grammars can tell you this.  So, for
% example, Bos et al (2004) would get bet(x,y) for "I bet 5 pounds",
% and bet(x,y,z) for "I bet you 5 pounds" and bet(x,p) for "I bet it's
% going to rain": I.e., the vocab of the semantic formalism is
% constructed automatically from NL word forms, and the predicate
% symbols in that automatically constructed vocab vary in (a) valency,
% and (b) the semantic dependencies represented by a given argument
% position (e.g., the second argument of "bet" in "I bet 5 pounds" is
% 5 pounds, and for "I bet it's raining" it's "it's raining".
% 
and long-distance dependencies.  
% TODO: search Steedman and Clark EMNLP 2004 for the performance
% figures on object relative clauses (which is constructions like
% "The man that I met", "the man that I think likes me", but not "the
% man that likes me"; they get about 50% of them exactly right; this
% is impressive but shows it's error prone.
Forcing a shallow parser
to make a decision about these phenomena therefore can be a major source
of errors.  It would be preferrable to give the parser the option to
declare itself not sufficiently informed to make the decision, and
to leave the decision open -- i.e., to compute a partial semantic
representation that neither over-determines nor under-determines the
semantic information revealed by the (shallow) syntactic analysis, and
to complete the semantic information later, via inferences that
exploit information extraneous to the shallow parser.

In this paper, we focus on one particular approach to representing
partial semantic information:
Robust Minimal Recursion Semantics ({\sc rmrs},
\cite{copestake:2003}).  {\sc rmrs} is a generalisation of {\sc mrs}
\cite{copestake:etal:2005}, designed to support underspecification both
of scope relations and of the predicate-argument structure.
\cite{copestake:2003, frank:2004} use it to
support semantic constructions by shallow parsers
ranging in depth from {\sc pos} taggers to chunk parsers  
and intermediate parsers such as {\sc rasp}
\cite{briscoe:etal:2006}.  {\sc mrs} analyses
derived from deep grammars, such as the English Resource Grammar
\cite{copestake:flickinger:2000} can also be seen as special cases of
{\sc rmrs} representations.

The key contribution we make is to cast {\sc rmrs}, for the first
time, as a logic with a clearly defined model theory.  This has a
number of advantages.  First, it allows us to formally relate \rmrs\ to
other framework such as \mrs\ and predicate logic in order to
establish the compatibility of \rmrs\ with the approaches to robust
semantic analysis mentioned above.  Second, it provides us with a
notion of logical entailment, which we can use to verify the
compatibility of \rmrs\ representations computed from different
parsers, such as a deep and a shallow parser.  Finally, it is only
through rigid formalisation that we can give users of \rmrs\ access to
inference systems and to algorithms for efficiently computing the
possible fully specific semantic representations described by a
partial \rmrs.

>>>>

Representing semantics as a logical form that supports automated
inference and model construction is vital for deeper language
engineering tasks, such as dialogue systems.  Such detailed semantic
representations can be obtained from hand-crafted deep grammars (e.g.,
\cite{butt:etal:1999,copestake:flickinger:2000}), but this
tends to lack robustness: hand-crafted grammars fail to cover all
words and linguistic constructions; and unedited text and speech
contains ill-formed phrases, which by design deep grammars do not
handle.

@@

However, there are certain semantic
phenomena that these robust
approaches don't capture reliably, including quantifier scope, optional
arguments,
% COMMENT: this issue about optional arguments is really too cryptic.
% Basically, there is no way for assessing by *any* learned language
% processor whether a single word form that appears in
% a number of different syntactic valencies is an example of lexical
% sense ambiguity, or they are all examples of the same sense being
% used in syntactically different ways (i.e., we have optional
% arguments).  Only hand-crafted grammars can tell you this.  So, for
% example, Bos et al (2004) would get bet(x,y) for "I bet 5 pounds",
% and bet(x,y,z) for "I bet you 5 pounds" and bet(x,p) for "I bet it's
% going to rain": I.e., the vocab of the semantic formalism is
% constructed automatically from NL word forms, and the predicate
% symbols in that automatically constructed vocab vary in (a) valency,
% and (b) the semantic dependencies represented by a given argument
% position (e.g., the second argument of "bet" in "I bet 5 pounds" is
% 5 pounds, and for "I bet it's raining" it's "it's raining".
% 
and long-distance dependencies (for instance, \cite{clark:etal:2004}
report that the parser used in \cite{bos:etal:2004} 
yields semantic dependencies from the unbounded
dependencies borne from
object extraction---e.g., {\em the man that I met\dots}---with 63\%
accuracy).
Forcing a shallow parser
to make a decision about these phenomena therefore can be a major source
of errors.  Depending on the application, it may be preferrable to
give the parser the option to 
declare itself not sufficiently informed to make the decision, and
to leave the decision open---i.e., to compute a partial semantic
representation that neither over-determines nor under-determines the
semantic information revealed by the (shallow) syntactic analysis, and
to complete the semantic information later, via inferences that
exploit information extraneous to the shallow parser.

In this paper, we focus on one particular approach to representing
partial semantic information:
Robust Minimal Recursion Semantics ({\sc rmrs},
\cite{copestake:2003}).  {\sc rmrs} is a generalisation of {\sc mrs}
\cite{copestake:etal:2005} that is designed to support underspecification both
of scope relations and of the predicate-argument structure.
\cite{copestake:2003,frank:2004} use it to
support semantic constructions by shallow parsers
ranging in depth from {\sc pos} taggers to chunk parsers  
and intermediate parsers such as {\sc rasp}
\cite{briscoe:etal:2006}.  {\sc mrs} analyses
derived from deep grammars, such as the English Resource Grammar
\cite{copestake:flickinger:2000} can also be seen as special cases of
{\sc rmrs} representations.

The key contribution we make is to cast {\sc rmrs}, for the first
time, as a logic with a clearly defined model theory.  This has a
number of advantages.  First, it allows us to formally relate \rmrs\ to
other frameworks such as \mrs\ and predicate logic in order to
establish the compatibility of \rmrs\ with the approaches to robust
semantic analysis mentioned above.  Second, it provides us with a
notion of logical entailment, which we can use to verify the
compatibility of \rmrs\ representations computed from different
parsers, such as a deep and a shallow parser.  Finally, it is only
through rigid formalisation that we can give users of \rmrs\ access to
inference systems and to algorithms for efficiently computing the
possible fully specific semantic representations described by the \rmrs
output of a shallow parser.
